News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-03-24 rupnik2015.tex
Update

\subsection{Evaluation Of Cross-Lingual Event Linking}

In order to determine how accurately we can predict cluster equivalence we have performed two experiments in multi-lingual setting using English, German and Spanish language for which we had labeled data to evaluate the linking performance. In the first experiment we wanted to test how well do the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation. In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning and testing process is repeated 10 times, each time with a different left-out fold.

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. To test which of these algorithms performs best we performed the following test. As a training set we used common Wikipedia alignment for all three languages. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA or the LSI algorithm, while $k$-means performs the worse. To decide which method to use we made some further experiments. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches on Wikipedia data by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. Similar performance of the LSI and CCA for cluster linking can be explained by domain missmatch, we train and test our model  AMRR performance on Wikipedia data but at the end we apply it to the news domain. Sadly we do not have any readily available data for this domain so we were not able to test properly.
Furthermore we inspect how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}. Large clusters aggregate a lot of information so the accuracy is not so important as long as we assign to correct clusters. In the case of small clusters accuracy should matter more.  As expected we obtained that it is harder to link small clusters and we need better retrieval accuracy in such cases. CCA outperforms other methods in this sense, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}. Added accuracy is also beneficial when we want
to pick good representatives of clusters. Therefore we decided to use CCA for cluster linking.

@@

\item \textbf{Features} - The experiments show that the hub CCA based features present good baseline, which can greatly benefit from additional semantic extraction based features. Even though the CCA based features do not greatly improve the system performance when the semantic features are used, there are two main benefits in the approach: the linking process can be sped up (smaller candidate cluster sets), and in the case when semantic extraction is not possible.

\item \textbf{Future work} - Currently the system is loosely coupled - the language component is built independently of the rest of the system, in particular the linking component. It is possible that better embeddings can be obtained by methods that jointly optimize a classification task and the embedding.

    Another point of interest is to evaluate the system on languages with scarce linguistic resources, where semantic annotation might not be available. For this reason, the labelled dataset of linked clusters should to be extended first. The mate retrieval evaluation showed that even for language pairs with no training set overlap, the hub cca recovers some signal.

@@

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) that produces $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for every language, for that we need 1.5 GB of memory, we store documents as 500 dimensional vectors. We must note that time complexity of similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article we compute top $10$  most similar ones in every other language.
For all linear algebra matrix and vector operations we use high performance numerical linear algebra libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500\times500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages.
 Usage of proxy matrices enables projection of all input documents in the common space and handling language pairs with missing or low alignment. That enables us to do block-wise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add new language we only need to compute a projector matrix and proxy matrices with all other already available languages.

Add the conclusion of the story and the wrap up, Andrej needs help with this.
We presented the system ...

>>>>

\subsection{Evaluation Of Cross-Lingual Event Linking}
In order to determine how accurately we can predict cluster equivalence we have performed two experiments in multi-lingual setting using English, German and Spanish language for which we had labeled data to evaluate the linking performance. In the first experiment we wanted to test how well do the individual approaches for cross-lingual article linking perform when used for linking the clusters about the same event. In the second experiment we tested how accurate is the prediction model when trained on different subsets of learning features. To evaluate the prediction accuracy for a given dataset we used 10-fold cross validation. In this technique, the dataset is partitioned into 10 folds, where 9 folds are used for training a model and the left-out fold is then used to test the accuracy of the model. The learning and testing process is repeated 10 times, each time with a different left-out fold.

In Section~\ref{sec:models} we have described three main algorithms for identifying similar articles in different languages. These algorithms were $k$-means, LSI and CCA. As a training set we used common Wikipedia alignment for all three languages. To test which of these algorithms performs best we performed the following test. Using each algorithm, we analyzed all articles in Event Registry to find the most similar articles in other languages. The values of the cross-lingual article linking features (\texttt{linkCount} and \texttt{avgSimScore}) were then recomputed for all annotated learning examples. A dataset with only these two features and the label was constructed and used to test how accurate prediction models can be induced using the data. The results of the experiment are shown in Table~\ref{table:linkingEvalAlgos}. The table shows for each of the algorithms the obtained classification accuracy, precision and recall. We can see that the most informative features for cluster linking are obtained using the CCA algorithm, while $k$-means performs the worse. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches by holding out $15000$ aligned test documents and using $300000$ aligned documents
Furthermore we inspected how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}.  As expected we obtained that it is harder to link small clusters and we need better retrieval. Hub CCA outperforms other methods in this sense also, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}.

To decide which method to use we made some further experiments. Firstly we verified the AMRR (average of mean reciprocal ranks over all language pairs) performance of approaches on Wikipedia data by holding out $15000$ aligned test documents and using $300000$ aligned documents
 as the training set. Figure~\ref{pic:AMRR} shows AMRR score as the function of the number of feature vectors. It is clear that CCA outperforms LSI approach and K-means lags far behind when testing on Wikipedia data. Similar performance of the LSI and CCA for cluster linking can be explained by domain missmatch, we train and test our model  AMRR performance on Wikipedia data but at the end we apply it to the news domain. Sadly we do not have any readily available data for this domain so we were not able to test properly.
Furthermore we inspect how number of features influences the accuracy of cluster linking. As we can see from Table~\ref{table:linkingEvalAlgos} choosing number of features vectors larger than $500$ barely effects linking performance. That is probably due to overfitting. Moreover we tested the performance of detecting large/small cluster linking, see Table~\ref{table:linkingEvalAlgosLargeSmall}. Large clusters aggregate a lot of information so the accuracy is not so important as long as we assign to correct clusters. In the case of small clusters accuracy should matter more.  As expected we obtained that it is harder to link small clusters and we need better retrieval accuracy in such cases. CCA outperforms other methods in this sense, probably as it has the best retrieval performance, see Figure~\ref{pic:AMRR}. Added accuracy is also beneficial when we want
to pick good representatives of clusters. Therefore we decided to use CCA for cluster linking.

@@

@@

Currently we compute similarities over $24$ languages with tags: eng, spa, deu, zho, ita, fra, rus, swe, nld, tur, jpn, por, ara, fin, ron, kor, hrv, tam, hun, slv, pol, srp, cat, ukr but we support up to $100$ top Wikipedis languages. Our data stream is Newsfeed (http://newsfeed.ijs.si/) that provides $430k$ unique articles per day. Our system currently computes 400k similiarites per 200ms, that means that we compute $16 \cdot 10^{10}$ similarities per day. We
store one day buffer for every language, for that we need 1.5 GB of memory, we store documents as 500 dimensional vectors. We must note that time complexity of similarity computations scales linearly with dimension of the feature space and does not  depend of number of languages. For each article we compute top $10$  most similar ones in every other language.
For all linear algebra matrix and vector operations we use high performance numerical linear algebra libraries as BLAS, OPENBLAS and Intel MKL. This currently allows us to process more than million articles per day.
In our current implementation we use variation of hub approach. Our projector matrices are of size $500\times 300000,$ so every projector takes about $1.1$ GB of RAM. Moreover we need proxy matrices of size $500\times500$ for every language pair. That is 0.5 GB for $24$ languages and $9.2$ GB for $100$ languages. All together we need around 135 GB of RAM for the system with 100 languages.
 Usage of proxy matrices enables projection of all input documents in the common space and handling language pairs with missing or low alignment. That enables us to do block-wise similarity computations and
additional improvement in our system efficiency. Our code can be therefore easily parallelized using matrix multiplication instead of doing more matrix - vector multiplications. This speeds up our code by factor around $4.$ In this way we obtain some caching gains and ability to use vectorization.
Our system is also easily extendable. If we want to add new language we only need to compute a projector matrix and proxy matrices with all other already available languages.
Add the conclusion of the story and the wrap up, Andrej needs help with this. We presented the system ...
